Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-03-23 paper-new/extensional-model.tex
more work on response

To additionally model the error terms, we add a requirement that there
is a natural transformation $\mho : 1 \to U$. The naturality
requirement encodes that strict morphisms (e.g., the denotations of
evaluation contexts) preserve errors.

We can then model CBV terms and types in a straightforward adaptation
of Levy's interpretation of CBV in CBPV. We interpret types $A$ as
objects $A \in \mathcal V$ and CBV terms $\Gamma \vdash M : A$ as
morphism of any of the equivalent forms $\mathcal E(F(\times\Gamma),
F(A)) \cong \mathcal V(\times\Gamma \vdash UF(A)) \cong \mathcal
E(F(1), \Gamma \to F(A))$. The most interesting type translation is
the CBV function type: $A \ra A' = U(A \to F A')$.
%
Such a model validates all type-based equational reasoning, i.e.,
$\beta\eta$ equality, and models the introduction and elimination
rules for CBV.
%
Thus a CBPV model is sufficient to interpret the CBV term language. We
will require additional structure to interpret the precision and type
casts.

>>>>

To additionally model the error terms, we add a requirement that there
is a natural transformation $\mho : 1 \to U B$. The naturality
requirement encodes that strict morphisms (e.g., the denotations of
evaluation contexts) preserve errors.

We can then model CBV terms and types in a straightforward adaptation
of Levy's interpretation of CBV in CBPV. We interpret types $A$ and
contexts $\Gamma$ as objects $A,\Gamma \in \mathcal V$ and CBV terms
$\Gamma \vdash M : A$ as morphisms of any of the equivalent forms
$\mathcal E(F\Gamma, FA) \cong \mathcal V(\Gamma \vdash UFA) \cong
\mathcal E(F(1), \Gamma \to F(A))$. The most interesting type
translation is the CBV function type: $A \ra A' = U(A \to F A')$.
%
Such a model validates all type-based equational reasoning, i.e.,
$\beta\eta$ equality, and models the introduction and elimination
rules for CBV.
%
Thus a CBPV model is sufficient to interpret the CBV term language. We
will require additional structure to interpret the precision and type
casts.
